Ordinal Models

### 2024-12-18 misc/discrep.qmd
update - marginal effects

>>>>

# s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
#   filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
#   filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
#   mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

# s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
#   filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
#   filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |>
#   mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

### 2024-12-18 misc/odds_ratio.qmd
update - marginal effects

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

@@

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

@@

ordinal_model_s2_logit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit)

pred_summary_s2 <- ordinal_model_s2_logit %>%
  epred_draws(newdata = s2_agg1, re_formula = NA,ndraws=200) %>%
  group_by("Reference Class"=refClass, rounded, "% Goal"=pct_goal, Category=.category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

# Create a new data frame with the predictor values
new_data <- expand.grid(
  refClass = c("kWh", "Percentage", "USD"),
  id = unique(s2_agg1$id)[1],      # Take first id
  state = unique(s2_agg1$state)[1], # Take first state,
  rounded = unique(s2_agg1$rounded)[1], # Take first rounded
  pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
)

@@

# Create a new data frame with the predictor values
# new_data <- expand.grid(
#   refClass = c("kWh", "Percentage", "USD"),
#   id = unique(s2_agg1$id)[1],      # Take first id
#   state = unique(s2_agg1$state)[1], # Take first state,
#   rounded = unique(s2_agg1$rounded)[1], # Take first rounded
#   pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
# )

@@

ordinal_model_s2_logit_int2 <- brm(
  accuracy_level ~ refClass +(rounded*pct_goal) + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_int2.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_int2)

@@

ordinal_model_s2_logit_int3 <- brm(
  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_int3)

@@

ordinal_model_s2_logit_c1rf_int3<- brm(
  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state) + (1|calc),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_c1rf_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_c1rf_int3)

@@

ordinal_model_s2_logit_c1add_int3<- brm(
  accuracy_level ~ (refClass * rounded*pct_goal) + calc + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.97), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_c1add_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_c1add_int3)

@@

@@

>>>>

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |>
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.05 ~ "0.01-5% error",
                    TRUE ~ "Over 5% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.05 ~ "0.01-5% error",
                TRUE ~ "Over 5% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", ".01-5% error", "Over 5% error"),
            ordered = TRUE
        ))

@@

s2_agg <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    close_match = abs(pct_change - pct) <= 0.02,
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.05 ~ "0.01-5% error",
                    TRUE ~ "Over 5% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match","0.01-5% error", "Over 5% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.05 ~ ".01-5% error",
                TRUE ~ "Over 5% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "01-5% error", "Over 5% error"),
            ordered = TRUE
        ))

@@

ordinal_model_s2_logit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit)

pred_summary_s2 <- ordinal_model_s2_logit %>%
  epred_draws(newdata = s2_agg, re_formula = NA,ndraws=200) %>%
  group_by("Reference Class"=refClass, rounded, "% Goal"=pct_goal, Category=.category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

# Create a new data frame with the predictor values
new_data <- expand.grid(
  refClass = c("kWh", "Percentage", "USD"),
  id = unique(s2_agg$id)[1],      # Take first id
  state = unique(s2_agg$state)[1], # Take first state,
  rounded = unique(s2_agg$rounded)[1], # Take first rounded
  pct_goal = unique(s2_agg$pct_goal)[1] # Take first pct_goal
)

@@

# Create a new data frame with the predictor values
# new_data <- expand.grid(
#   refClass = c("kWh", "Percentage", "USD"),
#   id = unique(s2_agg$id)[1],      # Take first id
#   state = unique(s2_agg$state)[1], # Take first state,
#   rounded = unique(s2_agg$rounded)[1], # Take first rounded
#   pct_goal = unique(s2_agg$pct_goal)[1] # Take first pct_goal
# )

@@

ordinal_model_s2_logit_int2 <- brm(
  accuracy_level ~ refClass +(rounded*pct_goal) + (1|id)+ (1|state),
  data = s2_agg,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_int2.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_int2)

@@

ordinal_model_s2_logit_int3 <- brm(
  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state),
  data = s2_agg,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_int3)

@@

ordinal_model_s2_logit_c1rf_int3<- brm(
  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state) + (1|calc),
  data = s2_agg,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_c1rf_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_c1rf_int3)

@@

ordinal_model_s2_logit_c1add_int3<- brm(
  accuracy_level ~ (refClass * rounded*pct_goal) + calc + (1|id)+ (1|state),
  data = s2_agg,
  family = cumulative("logit"),
  cores = 4,
  iter = 3000,
  control = list(adapt_delta = 0.97), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 3), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_c1add_int3.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit_c1add_int3)

@@

fixef(ordinal_model_s2_logit_c1add_int3)

library(marginaleffects)
library(tinytable)

ame <- avg_slopes(
  ordinal_model_s2_logit_c1add_int3, 
  variables = "refClass"
)
print(ame)

ame_by_rounded <- avg_slopes(
  ordinal_model_s2_logit_c1add_int3,
  variables = "refClass",
  by = c("rounded", "accuracy_level")
)
print(ame_by_rounded)

ame_by_rounded_pct_goal <- avg_slopes(
  ordinal_model_s2_logit_c1add_int3,
  variables = "refClass",
  by = c("rounded", "pct_goal", "accuracy_level")
)
print(ame_by_rounded_pct_goal)

conditional_slopes <- slopes(
    ordinal_model_s2_logit_c1add_int3,
    variables = "refClass",
    newdata = datagrid(
        rounded = c('Not Rounded', 'Rounded'),
        pct_goal = c("10%", "15%")
    )
)
conditional_slopes |> head(n=25)

plot_slopes(
    ordinal_model_s2_logit_c1add_int3,
    variables = "refClass",
    condition = c("rounded", "pct_goal")
)

avg_slopes(
    ordinal_model_s2_logit_c1add_int3,
    variables = "refClass",
    hypothesis = "b1 - b2 = 0"  # Compare effects for Rounded vs Not Rounded
)
#  Estimate   2.5 % 97.5 %
#    0.0569 -0.0528  0.159

# Term: b1-b2=0
# Type:  response 
# Columns: term, estimate, conf.low, conf.high 
avg_slopes(
    ordinal_model_s2_logit_c1add_int3,
    variables = "refClass",
    hypothesis = "b1 - b3 = 0"  # Compare effects for Rounded vs Not Rounded
)
#  Estimate  2.5 % 97.5 %
#   -0.0742 -0.192 0.0264

# Term: b1-b3=0
# Type:  response 
# Columns: term, estimate, conf.low, conf.high

avg_slopes(
ordinal_model_s2_logit_c1add_int3,
variables = "refClass",
by = c("rounded", "pct_goal"),
hypothesis = "(b1 - b2) - (b3 - b4) = 0"
)
#               Term Estimate   2.5 % 97.5 %
#  (b1-b2)-(b3-b4)=0   0.0201 -0.0797 0.0985

# Type:  response 
# Columns: term, estimate, conf.low, conf.high

# Example: Pairwise comparison of AMEs for refClass
pairwise_comparisons <- avg_slopes(
  ordinal_model_s2_logit_c1add_int3,
  variables = "refClass",
  hypothesis = "pairwise"
)
print(pairwise_comparisons)
#                                                                    Term Estimate    2.5 %    97.5 %
#  (Exact match, Percentage - kWh) - (Exact match, USD - kWh)                0.0575 -0.04893  0.170520
#  (Exact match, Percentage - kWh) - (0.01-2% error, Percentage - kWh)      -0.0738 -0.18531  0.028426
#  (Exact match, Percentage - kWh) - (0.01-2% error, USD - kWh)             -0.0618 -0.18001  0.051139
#  (Exact match, Percentage - kWh) - (2.01-15% error, Percentage - kWh)     -0.1011 -0.25716  0.035666
#  (Exact match, Percentage - kWh) - (2.01-15% error, USD - kWh)            -0.1122 -0.26659  0.023111
#  (Exact match, Percentage - kWh) - (Over 15% error, Percentage - kWh)     -0.1485 -0.34326  0.061167
#  (Exact match, Percentage - kWh) - (Over 15% error, USD - kWh)            -0.2098 -0.38566 -0.025617
#  (Exact match, USD - kWh) - (0.01-2% error, Percentage - kWh)             -0.1321 -0.25511 -0.019752
#  (Exact match, USD - kWh) - (0.01-2% error, USD - kWh)                    -0.1202 -0.23245 -0.018717
#  (Exact match, USD - kWh) - (2.01-15% error, Percentage - kWh)            -0.1615 -0.31554 -0.024206
#  (Exact match, USD - kWh) - (2.01-15% error, USD - kWh)                   -0.1707 -0.33209 -0.029627
#  (Exact match, USD - kWh) - (Over 15% error, Percentage - kWh)            -0.2082 -0.38108 -0.025534
#  (Exact match, USD - kWh) - (Over 15% error, USD - kWh)                   -0.2690 -0.47833 -0.045882
#  (0.01-2% error, Percentage - kWh) - (0.01-2% error, USD - kWh)            0.0108 -0.00632  0.037468
#  (0.01-2% error, Percentage - kWh) - (2.01-15% error, Percentage - kWh)   -0.0274 -0.07868  0.010135
#  (0.01-2% error, Percentage - kWh) - (2.01-15% error, USD - kWh)          -0.0381 -0.08937 -0.000373
#  (0.01-2% error, Percentage - kWh) - (Over 15% error, Percentage - kWh)   -0.0728 -0.16873  0.032944
#  (0.01-2% error, Percentage - kWh) - (Over 15% error, USD - kWh)          -0.1357 -0.22795 -0.024503
#  (0.01-2% error, USD - kWh) - (2.01-15% error, Percentage - kWh)          -0.0399 -0.09148 -0.002582
#  (0.01-2% error, USD - kWh) - (2.01-15% error, USD - kWh)                 -0.0495 -0.10824 -0.007293
#  (0.01-2% error, USD - kWh) - (Over 15% error, Percentage - kWh)          -0.0858 -0.17018  0.012997
#  (0.01-2% error, USD - kWh) - (Over 15% error, USD - kWh)                 -0.1463 -0.25832 -0.026034
#  (2.01-15% error, Percentage - kWh) - (2.01-15% error, USD - kWh)         -0.0080 -0.03343  0.004371
#  (2.01-15% error, Percentage - kWh) - (Over 15% error, Percentage - kWh)  -0.0420 -0.11437  0.024139
#  (2.01-15% error, Percentage - kWh) - (Over 15% error, USD - kWh)         -0.1032 -0.19343 -0.015885
#  (2.01-15% error, USD - kWh) - (Over 15% error, Percentage - kWh)         -0.0326 -0.11179  0.042446
#  (2.01-15% error, USD - kWh) - (Over 15% error, USD - kWh)                -0.0932 -0.17727 -0.012488
#  (Over 15% error, Percentage - kWh) - (Over 15% error, USD - kWh)         -0.0599 -0.18057  0.053139

emms_ref <- emmeans(ordinal_model_s2_logit_c1add_int3, ~ refClass, type = "link")
contrast_ref <- contrast(emms_ref, "trt.vs.ctrl", ref = "kWh") 
odds_ratios <- summary(contrast_ref, type = "response") # back-transform to odds
odds_ratios
#  contrast         odds.ratio lower.HPD upper.HPD
#  Percentage / kWh        2.9      0.25      10.3
#  USD / kWh               7.0      0.64      26.7

# Results are averaged over the levels of: rounded, pct_goal, calc 
# Point estimate displayed: median 
# Results are back-transformed from the log odds ratio scale 
# HPD interval probability: 0.95

comparisons <- hypothesis(
  ordinal_model_s2_logit_c1add_int3,
  hypothesis = c(
    "refClassPercentage - refClassUSD = 0",
    "refClassPercentage = 0",
    "refClassUSD = 0"
  ),
  class = "b"
)

pairs(emmeans(ordinal_model_s2_logit_c1add_int3, ~ refClass | rounded * pct_goal))

@@

```{markdown}

I need to generate and summarize publication quality group comparisons. Interpret the output from these initial analyses, and explain how I would report the effect of refClass in a manuscript, or if follow up analyses are necessary before we can confirm how best to report the effect. Also check the R code currently being used to ensure there are no mistakes. I mainly care about reporting the effect of refClass, but just don't want to make any mistakes in interpretation due to interactions. If additional analyses are necessary, provide R code for them. Also comment on if I should also assess group differences using odds ratios, and show R code to do this.

pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
library(marginaleffects)
library(tinytable)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

s2_agg <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    close_match = abs(pct_change - pct) <= 0.02,
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

ordinal_model_s2_logit_c1add_int3<- brm(
      accuracy_level ~ (refClass * rounded*pct_goal) + calc + (1|id)+ (1|state),
      data = s2_agg,
      family = cumulative("logit"),
      cores = 4,
      iter = 3000,
      control = list(adapt_delta = 0.97), # Recommended for ordinal models
      prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
                prior(normal(0, 3), class = "b")), # Priors for predictors
      file = paste0(here::here("data/model_cache",'s2_logit_c1add_int3.rds')) # Cache for efficiency
    )
    summary(ordinal_model_s2_logit_c1add_int3)
 Family: cumulative 
  Links: mu = logit; disc = identity 
Formula: accuracy_level ~ (refClass * rounded * pct_goal) + calc + (1 | id) + (1 | state) 
   Data: s2_agg (Number of observations: 784) 
  Draws: 4 chains, each with iter = 3000; warmup = 1500; thin = 1;
         total post-warmup draws = 6000

Multilevel Hyperparameters:
~id (Number of levels: 196) 
              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)     4.41      0.37     3.73     5.18 1.00     1382     2871

~state (Number of levels: 4) 
              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)     0.53      0.47     0.06     1.80 1.00     1046      864

Regression Coefficients:
                                              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept[1]                                     -1.80      0.80    -3.38    -0.27 1.00      814     1644
Intercept[2]                                     -0.26      0.79    -1.81     1.30 1.00      831     1824
Intercept[3]                                      3.59      0.80     2.03     5.17 1.00      854     1713
refClassPercentage                                1.15      0.91    -0.68     2.89 1.00      620     1179
refClassUSD                                       1.68      0.95    -0.19     3.55 1.00      725     1126
roundedRounded                                   -1.13      0.74    -2.60     0.30 1.00      763     1402
pct_goal15%                                      -0.89      0.74    -2.39     0.54 1.00      764     1342
calcNoCalculator                                  3.83      0.87     2.20     5.53 1.00      722     1685
refClassPercentage:roundedRounded                -0.35      0.99    -2.28     1.55 1.00      874     1723
refClassUSD:roundedRounded                        0.28      1.02    -1.74     2.24 1.00     1006     1762
refClassPercentage:pct_goal15%                   -0.10      0.98    -2.04     1.82 1.00      874     1755
refClassUSD:pct_goal15%                          -0.05      1.01    -2.02     1.91 1.00      994     1906
roundedRounded:pct_goal15%                        0.54      1.37    -2.07     3.26 1.00      698     1244
refClassPercentage:roundedRounded:pct_goal15%     0.60      1.83    -3.02     4.16 1.00      805     1526
refClassUSD:roundedRounded:pct_goal15%            0.62      1.88    -3.04     4.32 1.00      897     1565

Further Distributional Parameters:
     Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
disc     1.00      0.00     1.00     1.00   NA       NA       NA

Draws were sampled using sampling(NUTS). For each parameter, Bulk_ESS
and Tail_ESS are effective sample size measures, and Rhat is the potential
scale reduction factor on split chains (at convergence, Rhat = 1).

ame <- avg_slopes(
  ordinal_model_s2_logit_c1add_int3, 
  variables = "refClass"
)
print(ame)
          Group         Contrast Estimate    2.5 %   97.5 %
 Exact match    Percentage - kWh -0.08116 -0.19571  0.03058
 Exact match    USD - kWh        -0.14031 -0.25760 -0.02405
 0.01-2% error  Percentage - kWh -0.00626 -0.02044  0.00544
 0.01-2% error  USD - kWh        -0.01792 -0.03998 -0.00195
 2.01-15% error Percentage - kWh  0.02066 -0.00691  0.06562
 2.01-15% error USD - kWh         0.03031  0.00110  0.07913
 Over 15% error Percentage - kWh  0.06622 -0.03016  0.15240
 Over 15% error USD - kWh         0.12808  0.02140  0.22340

ame_by_rounded <- avg_slopes(
  ordinal_model_s2_logit_c1add_int3,
  variables = "refClass",
  by = c("rounded", "accuracy_level")
)
print(ame_by_rounded)

          Group     Term         Contrast     rounded Estimate     2.5 %    97.5 %
 Exact match    refClass Percentage - kWh Not Rounded -0.07940 -0.194010  0.031746
 Exact match    refClass USD - kWh        Not Rounded -0.11370 -0.231058  0.000454
 0.01-2% error  refClass Percentage - kWh Not Rounded -0.00865 -0.026939  0.006926
 0.01-2% error  refClass USD - kWh        Not Rounded -0.01731 -0.043440  0.000510
 2.01-15% error refClass Percentage - kWh Not Rounded  0.01113 -0.008613  0.055578
 2.01-15% error refClass USD - kWh        Not Rounded  0.01403 -0.007740  0.060557
 Over 15% error refClass Percentage - kWh Not Rounded  0.07524 -0.033826  0.173807
 Over 15% error refClass USD - kWh        Not Rounded  0.11672 -0.000602  0.219258
 Exact match    refClass Percentage - kWh Rounded     -0.08331 -0.208087  0.038742
 Exact match    refClass USD - kWh        Rounded     -0.16720 -0.295255 -0.040737
 0.01-2% error  refClass Percentage - kWh Rounded     -0.00371 -0.018239  0.008463
 0.01-2% error  refClass USD - kWh        Rounded     -0.01848 -0.041158 -0.001891
 2.01-15% error refClass Percentage - kWh Rounded      0.02904 -0.010886  0.083676
 2.01-15% error refClass USD - kWh        Rounded      0.04645  0.005512  0.101707
 Over 15% error refClass Percentage - kWh Rounded      0.05785 -0.033218  0.141438
 Over 15% error refClass USD - kWh        Rounded      0.13897  0.034533  0.238851

Type:  response 
Columns: term, group, contrast, rounded, estimate, conf.low, conf.high

ame_by_rounded_pct_goal <- avg_slopes(
  ordinal_model_s2_logit_c1add_int3,
  variables = "refClass",
  by = c("rounded", "pct_goal", "accuracy_level")
)

ame_by_rounded_pct_goal |> head(n=25)
